A Thee-Step Approach for Assessing Landscape Connectivity via Simulated Dispersal: African Wild Dog Case Study

### 2022-01-31 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, Lycaon pictus

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
% \newpage
% \begin{abstract}
% %------------------------------------------------------------------------------
% % Short Version
% %------------------------------------------------------------------------------
% \textbf{Short Version }
% The ability to disperse is contingent on a sufficient degree of landscape
% connectivity, which is why the identification and preservation of movement
% corridors that promote connectivity has become a task of extraordinary
% importance. Currently, ecologists rely on least-cost analysis or circuit theory
% to investigate connectivity, although both methods make several assumptions that
% are hardly met in reality. To address these issues, simulations from
% individual-based movement models have been proposed, yet a unified framework to
% simulate dispersal and quantify connectivity is lacking.
%
% Here, we propose a simple three-step approach that combines several
% already-existing methods to assess connectivity using simulated dispersal
% trajectories. In step one, we use integrated step selection functions to
% parametrize a mechanistic movement model rendering dispersal behavior. In step
% two, we apply the parametrized model as an individual-based movement model to
% simulate dispersal trajectories. In step three, we combine simulated
% trajectories into three complementary connectivity maps, each focusing on a
% different aspect of landscape connectivity.
%
% We showcase the application of the proposed approach using empirical data of
% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
% connectivity within the world's largest transboundary conservation area in
% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
% movement preferences, while also uncovering crucial dispersal corridors. With
% this analysis, we demonstrate that simulations from integrated step selection
% functions offer a simple, yet powerful alternative to traditional connectivity
% modeling techniques.
% \end{abstract}

@@

Here, we present a simple three-step approach to simulate dispersal movements
and to assess connectivity using empirical GPS movement data and environmental
habitat layers. In step one, we use integrated step selection functions to fit a
mechanistic movement model describing habitat and movement preferences of
dispersing individuals. In step two, we apply the parameterized movement model
to simulate dispersal trajectories. In step three, we derive three complementary
connectivity maps: a heatmap that highlights frequently traversed areas, a
betweenness map that pinpoints dispersal corridors, and a map of inter-patch
connectivity that indicates the presence and intensity of functional links
between habitat patches. As a case study, we applied the approach to GPS data
from dispersing individuals of the endangered African wild dog (\textit{Lycaon
pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA).

@@

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmap.png}
  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
  integrated step selection model that we fitted to the movement data of
  dispersing African wild dogs. To generate the heatmap, we rasterized and
  tallied all simulated trajectories. Consequently, the map highlights areas
  that are frequently traversed by virtual dispersers. For spatial reference we
  plotted a few selected NPs (dark gray). Additional heatmaps showing the
  traversal frequency when individuals move fewer than 2'000 steps are provided
  in Figure S4.}
  \label{Heatmap}
\end{figure}

>>>>

\textbf{Keywords:} dispersal, simulation, movement, integrated step-selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, Lycaon pictus

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
% \newpage
% \begin{abstract}
% %------------------------------------------------------------------------------
% % Short Version
% %------------------------------------------------------------------------------
% \textbf{Short Version }
% The ability to disperse is contingent on a sufficient degree of landscape
% connectivity, which is why the identification and preservation of movement
% corridors that promote connectivity has become a task of extraordinary
% importance. Currently, ecologists rely on least-cost analysis or circuit theory
% to investigate connectivity, although both methods make several assumptions that
% are hardly met in reality. To address these issues, simulations from
% individual-based movement models have been proposed, yet a unified framework to
% simulate dispersal and quantify connectivity is lacking.
%
% Here, we propose a simple three-step approach that combines several
% already-existing methods to assess connectivity using simulated dispersal
% trajectories. In step one, we use integrated step-selection functions to
% parametrize a mechanistic movement model rendering dispersal behavior. In step
% two, we apply the parametrized model as an individual-based movement model to
% simulate dispersal trajectories. In step three, we combine simulated
% trajectories into three complementary connectivity maps, each focusing on a
% different aspect of landscape connectivity.
%
% We showcase the application of the proposed approach using empirical data of
% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
% connectivity within the world's largest transboundary conservation area in
% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
% movement preferences, while also uncovering crucial dispersal corridors. With
% this analysis, we demonstrate that simulations from integrated step-selection
% functions offer a simple, yet powerful alternative to traditional connectivity
% modeling techniques.
% \end{abstract}

@@

Here, we present a simple three-step approach to simulate dispersal movements
and to assess connectivity using empirical GPS movement data and environmental
habitat layers. In step one, we use integrated step-selection functions to fit a
mechanistic movement model describing habitat and movement preferences of
dispersing individuals. In step two, we apply the parameterized movement model
to simulate dispersal trajectories. In step three, we derive three complementary
connectivity maps: a heatmap that highlights frequently traversed areas, a
betweenness map that pinpoints dispersal corridors, and a map of inter-patch
connectivity that indicates the presence and intensity of functional links
between habitat patches. As a case study, we applied the approach to GPS data
from dispersing individuals of the endangered African wild dog (\textit{Lycaon
pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA).

@@

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmap.png}
  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
  integrated step-selection model that we fitted to the movement data of
  dispersing African wild dogs. To generate the heatmap, we rasterized and
  tallied all simulated trajectories. Consequently, the map highlights areas
  that are frequently traversed by virtual dispersers. For spatial reference we
  plotted a few selected NPs (dark gray). Additional heatmaps showing the
  traversal frequency when individuals move fewer than 2'000 steps are provided
  in Figure S4.}
  \label{Heatmap}
\end{figure}
